Estimation of Intracortical Connectivity in a Dynamical Neural Field Model

### 2010-03-18 ltx/BrainIDE.tex
further updated paper following meeting on the 20/03/2010

where $f$ is spiking rate of populations $\mathbf{r}'$, $w$ is the spatial connectivity kernel, and $\Omega$ is the spatial domain, representing a cortical sheet or surface. The term $u\left(\mathbf{r},t\right)$ represents a known external input. The connectivity kernel is typically a Mexican hat shape described by the sum of weighted Gaussian functions. The shape of the kernel is not a critical parameter for our methodology so we shall not define it here to keep the derivation as general as possible.

@@

\section{Derivation of Finite Dimensional State-Space Model}
In order to implement standard estimation techniques we use a decomposition of the field using a set of Gaussian basis functions. Decomposition allows a continuous field to be represented by a finite dimensional state vector. This allows for the application of standard nonlinear, non-Gaussian state estimation methods such as the unscented Kalman filter. The field decomposition is described by
\begin{equation}\label{DefFieldDecomp}
	v_t\left(\mathbf{r}\right) \approx \boldsymbol{\phi}^{\top}\left(\mathbf{r}\right) \mathbf{x}_t
\end{equation}
where $\mathbf{\boldsymbol{\phi}}(\mathbf{r})$ is a vector of Gaussian basis functions that are scaled by the state vector, $\mathbf{x}_t$. As described in Eq.~\ref{Connectivity Kernel} the connectivity kernel can also be considered a decomposition into a set of basis functions in 
\begin{equation}\label{DefKernelDecomp}
	w\left(\mathbf{r}-\mathbf{r}'\right) = \boldsymbol{\theta}^{\top} \mathbf{\psi}\left(\mathbf{r}-\mathbf{r}'\right)
\end{equation}
We will assume we know the parametric form of the connectivity basis functions, where the parameter $\boldsymbol{\theta}$ is unknown. Each connectivity basis function can be considered a layer in the Wilson and Cowan model, representing short range excitation, surround inhibition and mid-range excitation. Making substitutions of~\ref{DefFieldDecomp} and~\ref{DefKernelDecomp} into~\ref{DiscreteTimeModel} we get
\begin{equation}\label{reduced continuous model}
	\boldsymbol{\phi}^{\top}\left( \mathbf{r} \right) \mathbf{x}_{t+1} = T\boldsymbol{\theta}^{\top}\int_\Omega  {\psi \left(\mathbf{r}-\mathbf{r}'\right) f\left( \boldsymbol{\phi} ^{\top}\left(\mathbf{r}'\right)\mathbf{x}_t \right)d\mathbf{r}'}  - \lambda\boldsymbol{\phi}^{\top}\left(\mathbf{r}\right)x_t + T u_t\left(\mathbf{r}\right) + e_t\left(\mathbf{r}\right).
\end{equation}
By defining the term
\begin{equation}\label{DefGamma}
	\boldsymbol{\Gamma}  = \int_\Omega  {\boldsymbol{\phi} \left(\mathbf{r}\right)\boldsymbol{\phi} ^{\top}\left(\mathbf{r}\right)d\mathbf{r}}
\end{equation}
and manipulating the model (see Appendix~\ref{Simplifying Decomposition}) we isolate the state vector to form a state-space model
\begin{eqnarray}\label{Homogeneous SS Model}
	\mathbf{x}_t &=& q(\mathbf{x}_t,\boldsymbol{\theta},\alpha) + k\left(u_t\left(\mathbf{r}\right)\right) + \mathbf{e}_t \\
	q(\mathbf{x}_t,\boldsymbol{\theta},\alpha) &=& T\boldsymbol{\Gamma}^{ - 1}\int_\Omega {\boldsymbol{\phi}\left(\mathbf{r}\right) \int_\Omega {f\left(\boldsymbol{\phi}^{\top}\left(\mathbf{r}'\right)\mathbf{x}_t\right)\psi^{\top} \left(\mathbf{r}-\mathbf{r}'\right)d\mathbf{r}'} d\mathbf{r}} \boldsymbol{\theta} - \lambda \mathbf{x}_t \\
	k(u_t(\mathbf{r})) &=& \boldsymbol{\Gamma}^{-1}T \int_\Omega{\boldsymbol{\phi} \left(\mathbf{r}\right) u_t\left(\mathbf{r}\right)d\mathbf{r}} \\
	\mathbf{e}_t &=& \boldsymbol{\Gamma}^{-1}\int_\Omega{\boldsymbol{\phi}\left(\mathbf{r}\right)e_t\left(\mathbf{r}\right)d\mathbf{r}} \\
\end{eqnarray}
where $\mathbf{e}_t$ is a zero-mean normal white noise term with the covariance (see Appendix D)
\begin{equation} 
\boldsymbol{\Sigma}_w=\boldsymbol{\Gamma}^{-1}\iint\limits_{\Omega}\boldsymbol{\phi}\left(\mathbf r\right) \rho\left(\mathbf{r}- \mathbf{r}' \right)\boldsymbol{\phi}\left(\mathbf{r}'\right)^{\top}d\mathbf{r}' d\mathbf{r}\boldsymbol{\Gamma}^{- \top} 
\end{equation}

@@

\section{Discrete Time Model}\label{Time Discretization}
To form the IDE neural field model a time discretization must be perform. We used a one-step Euler method where Eq.~\ref{FinalForm1} can be approximated by
\begin{equation}\label{Euler Approximation}	
\frac{v\left( r,t+T \right) - v\left( r,t\right)}{T} =   -\alpha v\left( r,t \right) + \int_\Omega  {w\left( |r-r'| \right)f\left( {v\left( r',t \right)} \right)dr'} + I\left(r,t\right).
\end{equation}
For clarity, we shall index time points in the discrete time form of the model using the subscript $t$ and the next time point as $t+1$. Rearranging Eq.~\ref{Euler Approximation} we get
\begin{eqnarray}\label{Euler Approximation}	
v_{t+1}\left( r\right) &=& v_t\left( r\right) -T \alpha v_t\left( r\right) + T \int_\Omega  {w\left( |r-r'| \right)f\left( {v_t\left( r'\right)} \right)dr'} + T I_t\left(r\right)
\end{eqnarray}
The discrete time form of the model is
\begin{equation}\label{Discrete Time Model1}
	v_{t+1}\left(r\right) = \lambda v_t\left(r\right) + T \int_\Omega { w\left(|r-r'|\right) f\left(v_t\left(r'\right)\right) dr'} + T I_t\left(r\right).
\end{equation}
where $\lambda = 1 - T \alpha$.

\section{Numerical Solution to IDE Model}\label{Space Discretization}
To solve the intregro-difference equation in Eq.~\ref{Discrete Time Model1} we define the spatial aspect of the model on a regular square $i,j$ grid of neural masses, where the spatial step size $\Delta \mathbf{r}_i = \Delta \mathbf{r}_j = \Delta $ giving
\begin{equation}\label{discrete space}
	v_{t+1}\left(r_{ij}\right) =  \lambda v_t(r_{ij}) + T \Delta^2 \sum_{i=1}^{i=I}{\sum_{j=1}^{j=J}{\tilde{w}\left( |r_{ij}-r_{ij}'| \right)f\left( v_t\left( r_{ij}'\right) \right)} } + T I_t\left(r\right) + \tilde{e}_t(r_{i,j})
\end{equation}
where $\tilde{e}_t(r_{i,j}) \sim \mathcal{N}\left(0,\Sigma\right)$ and $\tilde{w}(|r_{ij}-r_{ij}'|)$ discretized connectivity kernel described by
\begin{equation}\label{Discrete Connectivity Kernel}
	\tilde{w}(r_{ij},r_{ij}') = \boldsymbol{\theta}^{\top} e^{\frac{\left(\left(r_i'-r_i\right)\Delta\right)^2}{(2\sigma)^2} + \frac{\left(\left(r_j'-r_j\right)\Delta\right)^2}{(2\sigma)^2}}.
\end{equation}
The observation equation and kernel were discretized in the same way by
\begin{equation}\label{Discrete Observation Equation}
    y_k\left(r_{i_n j_n}\right) = \Delta^2\sum_{i=1}^{i=I}\sum_{j=1}^{j=J} \tilde{m}( |r_{i_n j_n}-r_{ij}'| ) v_k\left(r_{ij}'\right)  + \varepsilon_k( r_{i_n j_n})
\end{equation}
where
\begin{equation}\label{Discrete Observation Kernel}
	\tilde{m}(|r_{ij}-r_{ij}'|) = e^{\frac{\left(\left(r_i'-r_i\right)\Delta\right)^2}{(2\sigma)^2} + \frac{\left(\left(r_j'-r_j\right)\Delta\right)^2}{(2\sigma)^2}}.
\end{equation}
We have used periodic boundary conditions to alleviate problems associated with edge effects.

\section{Reduction to Finite State-Space Model}\label{Simplifying Decomposition}
Starting from Eq.~\ref{reduced continuous model} we multiply throughout by $\boldsymbol{\phi}(r)$ and integrate over the spatial domain $\Omega$ to get
\begin{equation}\label{Poo}
	\int_\Omega  {\boldsymbol{\phi} \left(r\right)\boldsymbol{\phi}^{\top}\left(r\right) dr} x_{t+1} = T \int_\Omega  {\boldsymbol{\phi} (r) \boldsymbol{\theta}^{\top} \int_\Omega  {\psi \left(|r-r'|\right)f\left(\boldsymbol{\phi}^{\top}\left(r'\right) x_t \right)dr'}dr}  - \lambda\int_\Omega {\boldsymbol{\phi}(r)\boldsymbol{\phi}^{\top}(r)dr} x_t + T \int_\Omega{\boldsymbol{\phi} \left(r\right) I_t\left(r\right)dr} + \int_\Omega{\boldsymbol{\phi} \left(r\right) e_t\left(r\right)dr}.
Substituting Eq.~\ref{DefGamma} into Eq.~\ref{Poo} and cross-multiplying by $\Gamma^{-1}$ gives 
\begin{equation}\label{Homogeneous SS Model}
	x_{t+1} = T\Gamma^{ - 1}\int_\Omega {\boldsymbol{\phi}\left(r\right) \int_\Omega {f\left(\boldsymbol{\phi}^{\top}\left(r'\right)x_t\right)\psi^{\top} \left(|r-r'|\right)dr'} dr} \boldsymbol{\theta} - \lambda x_t + \Gamma^{-1}T \int_\Omega{\boldsymbol{\phi} \left(r\right) u_t\left(r\right)dr} + \Gamma^{-1}\int_\Omega{\boldsymbol{\phi}\left(r\right)e_t\left(r\right)dr}.
The disturbance term can be rewritten as 
	\Gamma ^{ - 1}\int_\Omega  {\boldsymbol{\phi} \left(r\right)e_t\left(r\right)dr} = \int_\Omega{\boldsymbol{\phi}\left(r\right)\tilde{e}_t\left(r\right)dr},
where $\tilde{e}_t\left(r\right) \sim \mathcal{N}\left(0,R' \Sigma R\right)$ and $R$ is the Cholesky decomposition of $\Gamma^{-1}$. This provides the final simplified state-space model.
% \section{Ken's Discretization}
% Spatio-temporal interactions can be described by the continuous space continuous time stochastic differential equation
% \begin{equation}\label{differential eq}	
% \tau dv\left( r,t \right) = \left( -v\left( r,t \right) + \int_\Omega  {w\left( r,r' \right)f\left( {v\left( r',t \right)} \right)dr'} \right) dt  + d e\left( r,t \right).
% where $e\left( r,t \right)$ is a zero mean temporal Wiener process with incremental covariance $\Sigma_e(r)dt$. (not too sure about this noise since it needs to be white temporally but coloured spatially - which in continuous space-time I have never seen written ... has anyone?)
% To simplify the notation, but without loss of generality, \ref{differential eq} can be rewritten as
% \begin{equation}\label{differential simplified}
% 	dv\left( r,t \right) = q\left(v(r,t) \right) dt  + d e\left( r,t \right)
% Had $q\left(v(r,t) \right)$ been a linear function in $v$, then \ref{differential simplified} could be rewritten as
% \begin{equation}\label{differential linear}
% 	dv\left( r,t \right) = Av(r,t)dt  + d e\left( r,t \right)
% Considering temporal sampling instances $\delta$ apart given by $\{t_k = 0,~1,~2, \ldots$ \}, then a solution to \ref{differential linear} at each sampling time is given by
% \begin{equation}\label{linear solution}
% 	v(r,t_k+\delta) = \exp\{A(\delta)\}x(t) + \int_{t_k}^{t_k+\delta}\exp\{A(t_k+\delta-\tau)\}de(r,\tau)
% Let the discrete time random variable $e'(r,t_k)$ be
% \begin{equation}\label{discrete noise}
% 	e'(r,t_k) = \int_{t_k}^{t_k+\delta}\exp\{A(t_k+\delta-\tau)\}de(r,\tau)
% then, $e'(r,t_k)$ is a zero mean, temporally white noise term with covariance
% \begin{equation}\label{covariance of noise}
% 	E\left[e'(r,t_k),e'^\top(r,t_k) \right] = \int_t^{t+\delta}\exp\{A(t_k+\delta-\tau)\} \Sigma_e(r) \exp\{A(t+\delta-\tau)\} d\tau
% 
% But we are not so luck as to be able to go from \ref{differential simplified} to \ref{differential linear} and thus I doubt if we can obtain an exact solution to \ref{differential simplified} without an approximation. Agreed?
% 
% A possible approximate solution is the one you have already used. Thus, interpreting the difference equation \ref{differential simplified} as an integral equation gives
% \begin{equation}\label{euler}
% 	v\left(r,t\right) =  v(r,t_0) + \int_{t_0}^{t} q\left(v(r,\tau) \right) d\tau  + \int_{t_0}^{t} d e\left(r,\tau\right)
% Using an Euler approximation, \ref{euler} can be approximated at each sampling time by
% \begin{equation}\label{euler approx 1}
% 	v\left(r,t_{k+1} \right) =  v(r,t_k) + q\left(v(r,t_k)\right) (t_{k+1} - t_k) + (e(r,t_{k+1}) - e(r,t_k))
% which can be rewritten as
% \begin{equation}\label{euler approx 2}
% 	v\left(r,t_k + \delta \right) =  v(r,t_k) + \delta q\left(v(r,t_k)\right) + e'(r,t_k)
% where $e'(r,t_k)$ is a zero mean Gaussian random variable with covariance $\delta\Sigma_e$. (Note that, because of the way the noise is defined in \ref{differential eq} it is not $\delta^2$ but $\delta$)

@@

\begin{equation}
cov(e_t(\mathbf{r}),e_{t+\tau}(\mathbf{r'}))=
\begin{cases}
\beta\left(\mathbf{r}-\mathbf{r'}\right), & \tau=0 \\
0 & \mathrm{otherwise}
\end{cases}
\label{eq:FieldCovariance}
\end{equation}
The final state-space model is given by
\begin{equation}
x_{t + 1} = T q(x_t)\boldsymbol{\theta} -\lambda x_t + \Gamma^{-1}\int_\Omega  {\boldsymbol{\phi} ( \mathbf{r} )e_t( \mathbf{r} )d\mathbf{r}}
\end{equation}
then the covariance of the noise term is given by
\begin{eqnarray}
\mathbf{\Sigma}_w&=&\mathbf{\Gamma}^{-1}\mathbf E[\int_{\Omega}\boldsymbol{\phi}\left(\mathbf{r}\right)e_t\left(\mathbf{r}\right)d\mathbf{r} \int_{\Omega}\boldsymbol{\phi}\left(\mathbf{r}'\right)^{\top} e_t\left(\mathbf{r}'\right)d\mathbf{r}']\mathbf{\Gamma}^{- \top} \nonumber \\
&=&\mathbf{\Gamma}^{-1}\iint\limits_{\Omega} \boldsymbol{\phi}\left(\mathbf{r}\right) \mathbf E[e_t\left(\mathbf{r}\right)e_t\left(\mathbf{r}'\right)]\boldsymbol{\phi}\left(\mathbf{r}'\right)^{\top}d\mathbf{r}' d\mathbf r\mathbf{\Gamma}^{- \top} \nonumber\\
&=&\mathbf{\Gamma}^{-1}\iint\limits_{\Omega}\boldsymbol{\phi}\left(\mathbf r\right) \beta\left(\mathbf r- \mathbf r' \right)\boldsymbol{\phi}\left(\mathbf r'\right)^{\top}d\mathbf r' d\mathbf r\mathbf{\Gamma}^{- \top} 
\end{eqnarray}

>>>>

where $f$ is spiking rate of populations $\mathbf{r}'$, $w$ is the spatial connectivity kernel, and $\Omega$ is the spatial domain, representing a cortical sheet or surface. The term $u\left(\mathbf{r},t\right)$ represents a known external input. The connectivity kernel is typically a Mexican hat shape described by the sum of weighted Gaussian functions. The shape of the kernel is not a critical parameter for our methodology so we shall not define it until later to keep the derivation as general as possible.

@@

\section{Derivation of Finite Dimensional State-Space Model}
In order to implement standard estimation techniques we use a decomposition of the field using a set of Gaussian basis functions. Decomposition allows a continuous field to be represented by a finite dimensional state vector. This allows for the application of standard nonlinear, non-Gaussian state estimation methods such as the unscented Kalman filter. The field decomposition is described by
\begin{equation}\label{DefFieldDecomp}
	v_t\left(\mathbf{r}\right) \approx \boldsymbol{\phi}^{\top}\left(\mathbf{r}\right) \mathbf{x}_t
\end{equation}
where $\mathbf{\boldsymbol{\phi}}(\mathbf{r})$ is a vector of Gaussian basis functions that are scaled by the state vector, $\mathbf{x}_t$. The connectivity kernel can also be considered a decomposition into a set of basis functions in 
\begin{equation}\label{DefKernelDecomp}
	w\left(\mathbf{r}-\mathbf{r}'\right) = \boldsymbol{\theta}^{\top} \mathbf{\psi}\left(\mathbf{r}-\mathbf{r}'\right)
\end{equation}
We will assume we know the parametric form of the connectivity basis functions, where the parameter $\boldsymbol{\theta}$ is unknown. Each connectivity basis function can be considered a layer in the Wilson and Cowan model, representing short range excitation, surround inhibition and mid-range excitation. Making substitutions of~\ref{DefFieldDecomp} and~\ref{DefKernelDecomp} into~\ref{DiscreteTimeModel} we get
\begin{equation}\label{reduced continuous model}
	\boldsymbol{\phi}^{\top}\left( \mathbf{r} \right) \mathbf{x}_{t+1} = T\boldsymbol{\theta}^{\top}\int_\Omega  {\psi \left(\mathbf{r}-\mathbf{r}'\right) f\left( \boldsymbol{\phi} ^{\top}\left(\mathbf{r}'\right)\mathbf{x}_t \right)d\mathbf{r}'}  - \lambda\boldsymbol{\phi}^{\top}\left(\mathbf{r}\right)x_t + T u_t\left(\mathbf{r}\right) + e_t\left(\mathbf{r}\right).
\end{equation}
By defining the term
\begin{equation}\label{DefGamma}
	\boldsymbol{\Gamma}  = \int_\Omega  {\boldsymbol{\phi} \left(\mathbf{r}\right)\boldsymbol{\phi} ^{\top}\left(\mathbf{r}\right)d\mathbf{r}}
\end{equation}
and manipulating the model (see Appendix~\ref{Simplifying Decomposition}) we isolate the state vector to form a state-space model
\begin{eqnarray}\label{Homogeneous SS Model}
	\mathbf{x}_t &=& q(\mathbf{x}_t,\boldsymbol{\theta},\alpha) + k\left(u_t\left(\mathbf{r}\right)\right) + \mathbf{e}_t \\
	q(\mathbf{x}_t,\boldsymbol{\theta},\alpha) &=& T\boldsymbol{\Gamma}^{ - 1}\int_\Omega {\boldsymbol{\phi}\left(\mathbf{r}\right) \int_\Omega {f\left(\boldsymbol{\phi}^{\top}\left(\mathbf{r}'\right)\mathbf{x}_t\right)\psi^{\top} \left(\mathbf{r}-\mathbf{r}'\right)d\mathbf{r}'} d\mathbf{r}} \boldsymbol{\theta} - \lambda \mathbf{x}_t \\
	k(u_t(\mathbf{r})) &=& \boldsymbol{\Gamma}^{-1}T \int_\Omega{\boldsymbol{\phi} \left(\mathbf{r}\right) u_t\left(\mathbf{r}\right)d\mathbf{r}} \\
	\mathbf{e}_t &=& \boldsymbol{\Gamma}^{-1}\int_\Omega{\boldsymbol{\phi}\left(\mathbf{r}\right)e_t\left(\mathbf{r}\right)d\mathbf{r}} \\
\end{eqnarray}
where $\mathbf{e}_t$ is a zero-mean normal white noise term with the covariance (see Appendix D)
\begin{equation} 
\boldsymbol{\Sigma}_w=\boldsymbol{\Gamma}^{-1}\iint\limits_{\Omega}\boldsymbol{\phi}\left(\mathbf r\right) \rho\left(\mathbf{r}- \mathbf{r}' \right)\boldsymbol{\phi}\left(\mathbf{r}'\right)^{\top}d\mathbf{r}' d\mathbf{r}\boldsymbol{\Gamma}^{- \top} 
\end{equation}

@@

\section{Discrete Time Model}\label{Time Discretization}
To form the IDE neural field model a time discretization must be perform. We used a one-step Euler method where Eq.~\ref{FinalForm1} can be approximated by
\begin{equation}\label{Euler Approximation}	
\frac{v\left( \mathbf{r},t+T \right) - v\left( \mathbf{r},t\right)}{T} =   -\alpha v\left( \mathbf{r},t \right) + \int_\Omega  {w\left( \mathbf{r}-\mathbf{r}' \right)f\left( {v\left( \mathbf{r}',t \right)} \right)d\mathbf{r}'} + u\left(\mathbf{r},t\right).
\end{equation}
For clarity, we shall index time points in the discrete time form of the model using the subscript $t$ and the next time point as $t+1$. Rearranging Eq.~\ref{Euler Approximation} we get
\begin{eqnarray}\label{Euler Approximation}	
v_{t+1}\left( \mathbf{r}\right) &=& v_t\left( \mathbf{r}\right) -T \alpha v_t\left( \mathbf{r}\right) + T \int_\Omega  {w\left( \mathbf{r}-\mathbf{r}' \right)f\left( {v_t\left( \mathbf{r}'\right)} \right)d\mathbf{r}'} + T u_t\left(\mathbf{r}\right)
\end{eqnarray}
The discrete time form of the model is
\begin{equation}\label{Discrete Time Model1}
	v_{t+1}\left(\mathbf{r}\right) = \lambda v_t\left(\mathbf{r}\right) + T \int_\Omega { w\left(\mathbf{r}-\mathbf{r}'\right) f\left(v_t\left(\mathbf{r}'\right)\right) d\mathbf{r}'} + T u_t\left(\mathbf{r}\right).
\end{equation}
where $\lambda = 1 - T \alpha$.

\section{Numerical Solution to IDE Model}\label{Space Discretization}
To solve the intregro-difference equation in Eq.~\ref{Discrete Time Model1} we define the spatial aspect of the model on a regular square $i,j$ grid of neural masses, where the spatial step size $\Delta \mathbf{r}_i = \Delta \mathbf{r}_j = \Delta $ giving
\begin{equation}\label{discrete space}
	v_{t+1}\left(\mathbf{r}_{ij}\right) = \lambda v_t\left(\mathbf{r}_{ij}\right) + T \Delta^2 \sum_{i=1}^{i=I}{ \sum_{j=1}^{j=J}{ w\left( \mathbf{r}-\mathbf{r}_{ij}' \right)f\left( v_t\left( \mathbf{r}_{ij}'\right) \right)} } + T u_t\left(\mathbf{r}\right) + e_t(\mathbf{r}_{i,j})
\end{equation}
where $e_t(\mathbf{r}_{i,j}) \sim \mathcal{N}\left(0,\Sigma\right)$.
%  and $\tilde{w}(r_{ij}-r_{ij}'|)$ discretized connectivity kernel described by
% \begin{equation}\label{Discrete Connectivity Kernel}
% 	\tilde{w}(r_{ij},r_{ij}') = \boldsymbol{\theta}^{\top} e^{\frac{\left(\left(r_i'-r_i\right)\Delta\right)^2}{(2\sigma)^2} + \frac{\left(\left(r_j'-r_j\right)\Delta\right)^2}{(2\sigma)^2}}.
% \end{equation}
% The observation equation and kernel were discretized in the same way by
% \begin{equation}\label{Discrete Observation Equation}
%     y_k\left(r_{i_n j_n}\right) = \Delta^2\sum_{i=1}^{i=I}\sum_{j=1}^{j=J} \tilde{m}( |r_{i_n j_n}-r_{ij}'| ) v_k\left(r_{ij}'\right)  + \varepsilon_k( r_{i_n j_n})
% \end{equation}
% where
% \begin{equation}\label{Discrete Observation Kernel}
% 	\tilde{m}(|r_{ij}-r_{ij}'|) = e^{\frac{\left(\left(r_i'-r_i\right)\Delta\right)^2}{(2\sigma)^2} + \frac{\left(\left(r_j'-r_j\right)\Delta\right)^2}{(2\sigma)^2}}.
% \end{equation}
We have used periodic boundary conditions to alleviate problems associated with edge effects.

\section{Reduction to Finite State-Space Model}\label{Simplifying Decomposition}
Starting from Eq.~\ref{reduced continuous model} we multiply throughout by $\boldsymbol{\phi}(r)$ and integrate over the spatial domain $\Omega$ to get
\begin{eqnarray}\label{Poo}
	\int_\Omega  {\boldsymbol{\phi} \left(\mathbf{r}\right)\boldsymbol{\phi}^{\top}\left(\mathbf{r}\right) d\mathbf{r}} \mathbf{x}_{t+1} = T \int_\Omega  {\boldsymbol{\phi} (\mathbf{r}) \boldsymbol{\theta}^{\top} \int_\Omega  {\psi \left(\mathbf{r}-\mathbf{r}'\right)f\left(\boldsymbol{\phi}^{\top}\left(\mathbf{r}'\right) \mathbf{x}_t \right)d\mathbf{r}'}d\mathbf{r}} - \lambda\int_\Omega {\boldsymbol{\phi}(\mathbf{r})\boldsymbol{\phi}^{\top}(\mathbf{r})d\mathbf{r}} \mathbf{x}_t \\
	+ T \int_\Omega{\boldsymbol{\phi} \left(\mathbf{r}\right) u_t\left(\mathbf{r}\right)d\mathbf{r}} + \int_\Omega{\boldsymbol{\phi} \left(\mathbf{r}\right) e_t\left(\mathbf{r}\right)d\mathbf{r}}.
\end{eqnarray}
Substituting Eq.~\ref{DefGamma} into Eq.~\ref{Poo} and cross-multiplying by $\boldsymbol{\Gamma}^{-1}$ gives 
\begin{eqnarray}\label{Homogeneous SS Model}
	\mathbf{x}_{t+1} = T\boldsymbol{\Gamma}^{ - 1}\int_\Omega {\boldsymbol{\phi}\left(\mathbf{r}\right) \int_\Omega {f\left(\boldsymbol{\phi}^{\top}\left(\mathbf{r}'\right)\mathbf{x}_t\right)\psi^{\top} \left(\mathbf{r}-\mathbf{r}'\right)d\mathbf{r}'} d\mathbf{r}} \boldsymbol{\theta} - \lambda \mathbf{x}_t \\
	+ \boldsymbol{\Gamma}^{-1}T \int_\Omega{\boldsymbol{\phi} \left(\mathbf{r}\right) u_t\left(\mathbf{r}\right)d\mathbf{r}} + \boldsymbol{\Gamma}^{-1} \int_\Omega{\boldsymbol{\phi}\left(\mathbf{r}\right)e_t\left(\mathbf{r}\right)d\mathbf{r}}.
\end{eqnarray}

% \section{Ken's Discretization}
% Spatio-temporal interactions can be described by the continuous space continuous time stochastic differential equation
% \begin{equation}\label{differential eq}	
% \tau dv\left( r,t \right) = \left( -v\left( r,t \right) + \int_\Omega  {w\left( r,r' \right)f\left( {v\left( r',t \right)} \right)dr'} \right) dt  + d e\left( r,t \right).
% \end{equation}
% where $e\left( r,t \right)$ is a zero mean temporal Wiener process with incremental covariance $\Sigma_e(r)dt$. (not too sure about this noise since it needs to be white temporally but coloured spatially - which in continuous space-time I have never seen written ... has anyone?)
% To simplify the notation, but without loss of generality, \ref{differential eq} can be rewritten as
% \begin{equation}\label{differential simplified}
% 	dv\left( r,t \right) = q\left(v(r,t) \right) dt  + d e\left( r,t \right)
% \end{equation}
% Had $q\left(v(r,t) \right)$ been a linear function in $v$, then \ref{differential simplified} could be rewritten as
% \begin{equation}\label{differential linear}
% 	dv\left( r,t \right) = Av(r,t)dt  + d e\left( r,t \right)
% \end{equation}
% Considering temporal sampling instances $\delta$ apart given by $\{t_k = 0,~1,~2, \ldots$ \}, then a solution to \ref{differential linear} at each sampling time is given by
% \begin{equation}\label{linear solution}
% 	v(r,t_k+\delta) = \exp\{A(\delta)\}x(t) + \int_{t_k}^{t_k+\delta}\exp\{A(t_k+\delta-\tau)\}de(r,\tau)
% \end{equation}
% Let the discrete time random variable $e'(r,t_k)$ be
% \begin{equation}\label{discrete noise}
% 	e'(r,t_k) = \int_{t_k}^{t_k+\delta}\exp\{A(t_k+\delta-\tau)\}de(r,\tau)
% \end{equation}
% then, $e'(r,t_k)$ is a zero mean, temporally white noise term with covariance
% \begin{equation}\label{covariance of noise}
% 	E\left[e'(r,t_k),e'^\top(r,t_k) \right] = \int_t^{t+\delta}\exp\{A(t_k+\delta-\tau)\} \Sigma_e(r) \exp\{A(t+\delta-\tau)\} d\tau
% \end{equation}
% 
% But we are not so luck as to be able to go from \ref{differential simplified} to \ref{differential linear} and thus I doubt if we can obtain an exact solution to \ref{differential simplified} without an approximation. Agreed?
% 
% A possible approximate solution is the one you have already used. Thus, interpreting the difference equation \ref{differential simplified} as an integral equation gives
% \begin{equation}\label{euler}
% 	v\left(r,t\right) =  v(r,t_0) + \int_{t_0}^{t} q\left(v(r,\tau) \right) d\tau  + \int_{t_0}^{t} d e\left(r,\tau\right)
% \end{equation}
% Using an Euler approximation, \ref{euler} can be approximated at each sampling time by
% \begin{equation}\label{euler approx 1}
% 	v\left(r,t_{k+1} \right) =  v(r,t_k) + q\left(v(r,t_k)\right) (t_{k+1} - t_k) + (e(r,t_{k+1}) - e(r,t_k))
% \end{equation}
% which can be rewritten as
% \begin{equation}\label{euler approx 2}
% 	v\left(r,t_k + \delta \right) =  v(r,t_k) + \delta q\left(v(r,t_k)\right) + e'(r,t_k)
% \end{equation}
% where $e'(r,t_k)$ is a zero mean Gaussian random variable with covariance $\delta\Sigma_e$. (Note that, because of the way the noise is defined in \ref{differential eq} it is not $\delta^2$ but $\delta$)

@@

\begin{equation}
cov(e_t(\mathbf{r}),e_{t+\tau}(\mathbf{r'}))=
\begin{cases}
\beta\left(\mathbf{r}-\mathbf{r'}\right), & \tau=0 \\
0 & \mathrm{otherwise}
\end{cases}
\label{eq:FieldCovariance}
\end{equation}
The final state-space model is given by
\begin{equation}
\mathbf{x}_{t + 1} = T q(\mathbf{x}_t)\boldsymbol{\theta} -\lambda \mathbf{x}_t + \boldsymbol{\Gamma}^{-1}\int_\Omega  {\boldsymbol{\phi} ( \mathbf{r} )e_t( \mathbf{r} )d\mathbf{r}}
\end{equation}
then the covariance of the noise term is given by
\begin{eqnarray}
\mathbf{\Sigma}_w&=&\mathbf{\Gamma}^{-1}\mathbf E[\int_{\Omega}\boldsymbol{\phi}\left(\mathbf{r}\right)e_t\left(\mathbf{r}\right)d\mathbf{r} \int_{\Omega}\boldsymbol{\phi}\left(\mathbf{r}'\right)^{\top} e_t\left(\mathbf{r}'\right)d\mathbf{r}']\mathbf{\Gamma}^{- \top} \nonumber \\
&=&\mathbf{\Gamma}^{-1}\iint\limits_{\Omega} \boldsymbol{\phi}\left(\mathbf{r}\right) \mathbf E[e_t\left(\mathbf{r}\right)e_t\left(\mathbf{r}'\right)]\boldsymbol{\phi}\left(\mathbf{r}'\right)^{\top}d\mathbf{r}' d\mathbf r\mathbf{\Gamma}^{- \top} \nonumber\\
&=&\mathbf{\Gamma}^{-1}\iint\limits_{\Omega}\boldsymbol{\phi}\left(\mathbf r\right) \beta\left(\mathbf r- \mathbf r' \right)\boldsymbol{\phi}\left(\mathbf r'\right)^{\top}d\mathbf r' d\mathbf r\mathbf{\Gamma}^{- \top} 
\end{eqnarray}
